---
title: "Homework_04"
author: "Molly Ratliff"
date: "2024-02-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**1) Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the value of the variable z and print the value stored in z.**
```{r}
x <- 1.1
a <- 2.2
b <- 3.3
```
**a.**
```{r}
z <- ((x)^a)^b
print(z)
```
**b.**
```{r}
z <- (x^a)^b
print(z)
```
**c.**
```{r}
z <- 3*((x)^3) + 2*((x)^2) + 1
print(z)
```

**2) Using the rep and seq functions, create the following vectors:**

**a. (1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)**
```{r}
vec_a <- c(seq(from=1, to=8), seq(from=7, to=1))
print(vec_a)
```

**b. (1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)**
```{r}
vec_b <- c(1,2,3,4,5)
rep(x=vec_b, times=vec_b)
```

**c. (5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)**
```{r}
vec_c <- c(5,4,3,2,1)
rep(x=vec_c, times=vec_b)
```

**3) Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as x and y coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R’s inverse trigonometry functions (asin(), acos(), or atan()), convert these numbers into polar coordinates.**
```{r}
coords <- runif(2)
x <- coords[1]
y <- coords[2]
radius <- sqrt(x^2 + y^2)
theta <- atan(y/x)
pol_coords <- c(radius, theta)
print(pol_coords)
```

**4) Create a vector queue <- c("sheep", "fox", "owl", "ant") where queue represents the animals that are lined up to enter Noah’s Ark, with the sheep at the front of the line. Using R expressions, update queue as: **
```{r}
queue <- c("sheep", "fox", "owl", "ant")
print(queue)
```
**a. the serpent arrives and gets in line**
```{r}
queue <- c(queue, "serpent")
print(queue)
```
**b. the sheep enters the ark**
```{r}
queue <- queue[2:5]
print(queue)
```
**c. the donkey arrives and talks his way to the front of the line**
```{r}
queue <- c("doneky", queue)
print(queue)
```
**d. the serpent gets impatient and leaves**
```{r}
queue <- queue[1:4]
print(queue)
```
**e. the owl gets bored and leaves**
```{r}
queue <- queue[-3]
print(queue)
```
**f. the aphid arrives and the ant invites him to cut in line**
```{r}
queue <- append(queue, "aphid", 2)
print(queue)
```

**g. Finally, determine the position of the aphid in the line**
```{r}
which(queue == "aphid")
```

**5) Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7.**
```{r}
vec <- 1:100
vec <- vec[!(vec %% 2 == 0 | vec %% 3 == 0 | vec %% 7 == 0)]
print(vec)
```